simplified_likelihood

### 2023-05-22 inst/simplified_likelihood.Rmd
correct some stupid mistakes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

Observing that the fifth line may be neglected, R idiom for this would be as follows

```{r}
library("hyper2")
simplike <- function(n1,n2,n){
  H <- hyper3()
  H[c(p =1)] <- n2-2
  H[c(p1=1)] <- 1
  H[c(p2=1)] <- 1
  for(i in (n-n1+1):n)     { H[c(p1=1,p2=1,p=i)]  %<>% dec }
  for(i in (n-n2+2):(n-n1+1)){ H[c(p2=1,   p=i)]  %<>% dec }
  return(H)
} 
simplike(2,6,11)
simplike(1,6,11)
simplike(1,2,11)
```

>>>>

```{r setup, include=FALSE}
{ # this line vandalises this file: nonfinishers are ignored and should not be!
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

@@

```{r}
named_drivers <- function(n,...){ # twodrivers(20, ham=3, bottas=5)  # NB repeats allowed!
   a <- list(...)
   stopifnot(all(table(as.vector(unlist(a)))==1)) # stops two players having the same result
   s <- rep("other",n)
   for(i in seq_along(a)){
     s[a[[i]]] <- names(a)[i]   # ham=3 -> s[3] <- "ham"
   }
   return(ordervec2supp3(s))  
}
```

```{r}
(H <- named_drivers(20,ham=3,bott=5))
```

Thus `H` is a likelihood function on the assumption that the other
drivers are all clones of identical strength.  Maximizing this:

```{r readtable}
a <- read.table("formula1_2019.txt")[1:2,1:21]
a
f <- function(i){
  jj <- as.list(a[,i])
  suppressWarnings(jj <- as.numeric(jj))
  names(jj) <- rownames(a)
  jj[!is.na(jj)]   # NA entries ignored - THESE SHOULD BE NONFINISHERS
}

H <- hyper3()
for(i in seq_len(ncol(a))){
  jj <- f(i)
  if(all(!is.na(as.numeric(unlist(jj))))){
  }
}
H
```

```{r maxh,cache=TRUE}
maxp(H)
```
Thus we narrowly fail to reject the null at $\alpha = 0.05$.  Now
compare with the actual 2019 season:

```{r actual2019,cache=TRUE}
F1_2019 <- ordertable2supp(read.table("formula1_2019.txt",header=TRUE)[,1:21])
samep.test(F1_2019,c("Hamilton","Bottas"))
```

We see a much smaller p-value when using the complete dataset.  It is
not clear whether we would expect a larger or a smaller p-value here,
but it is not surprising to see a difference.  On the simplified case,
the estimated strength of `other` is about 0.016 (under the
alternative); but in the full case, most of the other drivers have
strength considerably more thn this.  I do not have a good way of
thinking about this at this time.  One other feature is that the
simplified likelihood results in different MLEs:

* simplified: $\widehat{p_\mbox{Ham}} = 0.49$, $\widehat{p_\mbox{Bot}}=0.31$
* full: $\widehat{p_\mbox{Ham}} = 0.38$, $\widehat{p_\mbox{Bot}}=0.13$

Consider the log-contrasts, $\log(p_1/p_2)$ giving
